Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
example

Although limiting the size of a handshake message seems an elementary problem to
repair, the example shows that there was a significant amount of changes
performed in the codebase which produced a negative impact in the project
maintainability. Even though the developer tries to simplify the codebase with
the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len} (~\ref{lst:func1}), it
also disrupts $2$ of the guidelines proposed by the Software Improvement Group
(SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a
new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit
(~\ref{lst:func5}) increasing the cyclomatic complexity and breaking the
\emph{Write Simple Units of Code} guideline. Moreover, it increases the codebase
size in $4$ lines of code ($13$ lines added, $9$ lines deleted) which contributes
to a potential violation of the
\emph{Keep Your Codebase Small} guideline. Hence, we have observed that the
impact of this refactoring on the OpenSSL maintainability was negative.

@@

There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a cross-site scripting vulnerability with an \texttt{escape}
function. Knowing which patterns are more likely to increase maintainability
issues is one step forward to bring awareness to security engineers of what
patterns need more attention. The taxonomy of security patterns used to answer
this question is the same as the one provided by the authors of the dataset~\cite{Reis:2017:IJSSE}.
Maintainability is measured separately for each pattern.
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed this section. The methodology
comprehends the following steps, as illustrated in Figure~\ref{fig:met}:
\begin{enumerate}
	\item Extract relevant data from a dataset containing
$716$ security refactorings collected from open-source software available on
GitHub~\cite{Reis:2017:IJSSE}.
	\item A baseline of regular commits was
randomly collected from the list of projects of the main dataset to evaluate the impact of regular commits on the maintainability of open-source
software.
\item Use the Software Improvement Group's\footnote{SIG's website is available
at https://www.sig.eu/ (Accessed on \today{})} web-based source code analysis
service \emph{Better Code Hub} (BCH)\footnote{BCH's website is available at
https://bettercodehub.com/ (Accessed on \today{})} to quantify
maintainability for both security and regular commits.
\end{enumerate}

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}
\Luis{Nao uses numeros no inico da frase. --> We found that 40.2\% of the security refactorings}

In particular, 40.2\% of the security refactoring have a negative impact on the
maintainability of open-source software, Figure~\ref{fig:secvsreg}. This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of safety. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and user-friendly~\cite{5287006, 7676144}. Unfortunately,
	there are several vulnerabilities that stay undisclosed. Thus, developers should not
	trust them blindly and check defaults.
	\item \textbf{Documentation} of software should provide developers with the best practices to implement safe and maintainable software. Unmaintainable code is difficult to test, analyze and re-use.
	\item \textbf{Proper Authorization Mechanisms} should be designed and implemented. This is
	one of the most complex security features to implement. Some mechanisms might assist authorization maintenance in long run, such as the combination of \texttt{HTTP} filters with \texttt{Spring Security} annotations.
	\item \textbf{Deprecated code} should be avoided by developers. They should integrate tools such as BCH on their projects CI/CD pipelines to help them be more aware of the maintainability of their code and issues (e.g., dead and duplicated code).
	\item\textbf{Programming Languages} should provide new mechanisms to easily implement security patterns without endangering maintainability.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Know Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
type of vulnerabilities are one line fixes --- adding an \texttt{escape}
function or changing the version of the component, respectively. Whereas,
Broken Authentication and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%

\Luis{Better Code Hub -> BCH}
The maintainability formula was inferred based on the reports of the Better Code Hub.
The high amount of different projects and backgrounds may require different maintainability
standards. However, BCH does have a representative benchmark of closed and open-source
software projects that is under the calculation of the thresholds used in the computation
of each maintainability guideline (Table~\ref{tab:guidelines})~\cite{Visser:2016:OREILLY,
Baggen2012}.

@@

\Luis{maintainability still may be affected -> maintainability may sitll be affected}

@@

\Luis{A primeira threat que falas é que os resultados podem nao generalizar
para software fechado. Eu acho que deviamos ser mais diretos. Para alem disso
tirava o "In contrast to the BCH benchmark,"}

In contrast to the BCH benchmark, this study uses a dataset that only covers
open-source software to perform the maintainability evaluation. However, BCH
thresholds are calculated from a set of both open and closed software.
Different programming languages may require different coding practices to
address software safety. The dataset comprises more commits in C and PHP, i.e.,
the dataset may not be properly representative of the population regarding to
programming languages.

@@

\Luis{pode ser uma cena só minha mas acho que fica melhor em vez de comecar
fraseds com "And," usar outros conetores tipo "Besides," "In addition,", etc. œ}

@@

\Luis{eu nao acho que usar 10 guidelines proposed by the Software Improvement
Group (SIG) seja algo que nos esteja a diferenciar. Acho que o a prinicpal
diferenca é que estamos a estudar maintainability issues noutro contexto.}

@@

\Luis{ tambem fizemos analise por linguagem, nesta ultima citacao deviamos falar sobre isso tb.}

@@

\Luis{secure security frameworks, nao percebi bem o que se quer dizer aí.}

@@

\Luis{many directions such as -> many directions:}
\Luis{evaluate the impact of programming languages of the
maintainability of each refactoring pattern; - > evaluate the impact of each refactoring pattern on the maintainbility of each  }

>>>>

Although limiting the size of a handshake message seems an elementary problem to
repair, the example shows that there was a significant amount of changes
performed in the codebase which produced a negative impact in the project
maintainability. Even though the developer tries to simplify the codebase with
the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len} (\ref{lst:func1}), it
also disrupts $2$ of the guidelines proposed by the Software Improvement Group
(SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a
new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit
(\ref{lst:func5}) increasing the cyclomatic complexity and breaking the
\emph{Write Simple Units of Code} guideline. Moreover, it increases the codebase
size in $4$ lines of code ($13$ lines added, $9$ lines deleted) which contributes
to a potential violation of the
\emph{Keep Your Codebase Small} guideline. Hence, we have observed that the
impact of this refactoring on the OpenSSL maintainability was negative.

@@

There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a cross-site scripting vulnerability with simple function to sanitize data. Knowing which patterns are more likely to increase maintainability
issues is one step forward to bring awareness to security engineers of what
patterns need more attention. The taxonomy of security patterns used to answer
this question is the same as the one provided by the authors of the dataset~\cite{Reis:2017:IJSSE}.
Maintainability is measured separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by nextcloud/server developers to a \\Cross-Site Scripting vulnerability},label={lst:fix},language=php]
 <ul>
    <li class='error'>
       <?php echo $l->t( 'Cloud not found' ); ?><br/>
-      <p class='hint'><?php if(isset($_['file'])) echo $_['file']?></p>
+      <p class='hint'><?php if(isset($_['file'])) echo htmlentities($_['file'])?></p>
    </li>
 </ul>

\end{lstlisting}
%
\section{Methodology}\label{sec:methodology}
%
The methodology used to measure the impact of security refactorings on the
maintainability of open-source software is discussed this section. The methodology
comprehends the following steps, as illustrated in Figure~\ref{fig:met}:
\begin{enumerate}
	\item Extract relevant data from a dataset containing
$716$ security refactorings collected from open-source software available on
GitHub~\cite{Reis:2017:IJSSE}.
	\item A baseline of regular commits was
randomly collected from the list of projects of the main dataset to evaluate the impact of regular commits on the maintainability of open-source
software.
\item Use the Software Improvement Group's\footnote{SIG's website is available
at https://www.sig.eu/ (Accessed on \today{})} web-based source code analysis
service \emph{Better Code Hub} (BCH)\footnote{BCH's website is available at
https://bettercodehub.com/ (Accessed on \today{})} to quantify
maintainability for both security and regular commits.
\end{enumerate}

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}

40.2\% of the security refactoring have a negative impact on the maintainability
of open-source software, Figure~\ref{fig:secvsreg}. This raises a new tradeoff when
developers need to fix vulnerabilities on their projects because they might be affecting
the software maintainability or even introducing new vulnerabilities. This study exhibits
proof that developers may have to reduce maintainability for the sake of safety. We understand
that developers should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and user-friendly~\cite{5287006, 7676144}. Unfortunately,
	there are several vulnerabilities that stay undisclosed. Thus, developers should not
	trust them blindly and check defaults.
	\item \textbf{Documentation} of software should provide developers with the best practices to implement safe and maintainable software. Unmaintainable code is difficult to test, analyze and re-use.
	\item \textbf{Proper Authorization Mechanisms} should be designed and implemented. This is
	one of the most complex security features to implement. Some mechanisms might assist authorization maintenance in long run, such as the combination of \texttt{HTTP} filters with \texttt{Spring Security} annotations.
	\item \textbf{Deprecated code} should be avoided by developers. They should integrate tools such as BCH on their projects CI/CD pipelines to help them be more aware of the maintainability of their code and issues (e.g., dead and duplicated code).
	\item\textbf{Programming Languages} should provide new mechanisms to easily implement security patterns without endangering maintainability.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring adds complexity to the codebase it is probably affecting the software maintainability. One evidence of that is the fact of \emph{Cross-Site Scripting} and \emph{Using Components with Know Vulnerabilities} refactorings endure more cases without any impact in the maintainability. Typically, these type of vulnerabilities are one line fixes --- adding an \texttt{escape} function or changing the version of the component, respectively. Whereas, Broken Authentication and vulnerabilities such as the one presented in Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to fix (several code lines deleted and added). Another example is the Apple authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872 details available at https://support.apple.com/en-us/HT208315 (Accessed on \today{})}, where any user could login as root with an empty password. Patrick Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on \today{})} and conclude that the flaw was due to the introduction of high cyclomatic complexity in a method used to verify the password. In sum, \emph{Cross-Site Scripting} and \emph{Using Components with Known Vulnerabilities} should take part of the developers preocupations along with the other patterns that explicitly hinder maintainability because despite their low complexity these are patterns that still produce a considerable negative impact on the maintainability. Real efforts have been made in creating and incorporating coding standards and best practices in software production. However, security is still far from perfection and fixing vulnerabilities might hinder software maintainability. Thus, it is important to provide tool support to developers to help them applying these patterns without harming software maintainability.

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the reports of the Better Code Hub.
The high amount of different projects and backgrounds may require different maintainablity
standards. However, BCH does have a representative benchmark of closed and open-source
software projects that is under the calculation of the thresholds used in the computation
of each maintainability guideline (Table~\ref{tab:guidelines})~\cite{Visser:2016:OREILLY,
Baggen2012}.

@@

@@

In contrast to the BCH benchmark, this study uses a dataset that only covers open-source software to perform the maintanibility evaluation. However, BCH thresholds are calculated from a set of both open and closed sotfware. Different programming languages may require different coding practices to address software safety. The dataset
comprises more commits in C and PHP, i.e., the dataset may not be properly representative of the population regarding to programming languages.

@@

@@

@@

@@

@@
